Canonical rings of $\bq$-divisors on Minimal Rational Surfaces

### 2015-08-04 surfaces-write-up/surface-write-up.tex
some clarifying edits in the pf of proj-eff, respond to one of Peter's todos

\begin{defn}
\label{defn:vec-sum}
For $\vec{v} \in \bq^n$, denote $\deg \vec{v} := \sum_{i = 0}
^n v_i$.
Let 
\begin{align*}
	\mss_i := \left \{\vec{v} \in \bz_{\geq 0}^{m + 1} \; : \;
\deg \vec v = c_i \right\}.	
\end{align*}
\end{defn}

@@

\noindent
has a minimal presentation over $R_{D'}$ consisting of the $\sum_{i = 0}^{r}
{{m + c_i - 1} \choose {c_i}}$ generators $F_i^{\vec{v}} := \frac{u^{d_i}
x^{\vec{v}}}{x_k^{c_i}}$ where $1 \leq i \leq r$, $\vec{v} \in \bz_{\geq 0}^{m + 1}$
with $\deg \vec v = c_i$. Furthermore, it has
relations generated by the following two forms:
\begin{enumerate}
	\item For each $(i, j)$ with $j \geq i + 2$ and each $\vec{v} \in \mss_i,
\vec{w} \in \mss_j$, there is a relation of the form: either
\begin{flalign*}
	&G_{i, j}^{\vec{v}, \vec{w}} = F_i^{\vec{v}} F_j^{\vec{w}}
	- \prod_{\vec{y} \in \mss_{h_{i, j}}} (F_{h_{i, j}}^{\vec{y}})
	^{g_{\vec{y}}} &(i < h_{i, j} < j) \\
	\text{or} \\
	&G_{i, j}^{\vec{v}, \vec{w}} = F_i^{\vec{v}} F_j^{\vec{w}}
	- \prod_{\vec{y} \in \mss_{h_{i, j}}} (F_{h_{i, j}}^{\vec{y}})^{g_{\vec{y}}}
	\cdot \prod_{ \vec{z} \in
	\mss_{h_{i, j} + 1}} (F_{k_{i, j} + 1}^{\vec{z}})^{g'_{\vec{z}}}
	&(i < h_{i, j} < h_{i, j} + 1 < j).
\end{flalign*} \todo{Peter: Can you clarify when these two cases occur: i.e. does the first only occur if $j = h_{i,j}+1$; otherwise, in cases when $i< h_{i,j} < h_{i,j} + 1< j$, do we get relations of both forms?}
	\item For each $(i, j)$ with
$j = i$ or $j = i + 1$ and each $\vec{v} \in \mss_i, \vec{w} \in
\mss_j$ with $\vec{v} \not\prec \vec{w}$ (see Definition
~\ref{defn:vec-order}) there is a relation of the form
\begin{align*}
	&L_{i, j}^{\vec{v}, \vec{w}} = F_i^{\vec{v}} F_j^{\vec{w}}
	- F_i^{\vec{y}} F_j^{\vec{z}} \\
\end{align*}

@@

Next, we show that the relations given in the statement of the theorem generate all relations. 
In particular, if $j \geq i + 2$ then $F_i^{\vec{v}} F_j^{\vec{w}}$
has a decomposition of the form ~\eqref{eqn:one-point-canonical-form} where $h$
depends on $i, j$, so we notate $h = h_{i, j} \in \{1, \ldots, r\}$. We also
have
that $i \leq h_{i, j} < j$ since $(d_i + d_j, c_i + c_j)$ is in the
$\bz_{\geq 0}$-span of $(d_{h_{i, j}}, c_{h_{i, j}})$ and
$(d_{h_{i, j} + 1}, c_{h_{i, j} + 1})$. Furthermore, we have that
$h_{i, j} \neq i, j-1$ as can be shown in the same manner as the analogous statement in O'Dorney
\cite[Theorem 6]{dorney:canonical}. This gives the relations
$G_{i, j}^{\vec{v}, \vec{w}}$.

>>>>

\begin{defn}
\label{defn:vec-sum}
For $\vec{v} \in \bq^n$, denote $\deg \vec{v} := \sum_{i = 0}
^n v_i$.
Let 
\begin{align*}
	\mss_i := \left \{\vec{v} \in \bz_{\geq 0}^{m + 1} \; : \;
\deg \vec v = c_i \right\}.	
\end{align*}

\noindent
be the set of vectors of degree $c_i$.
\end{defn}

@@

\noindent
has a minimal presentation over $R_{D'}$ consisting of the $\sum_{i = 0}^{r}
{{m + c_i - 1} \choose {c_i}}$ generators $F_i^{\vec{v}} := \frac{u^{d_i}
x^{\vec{v}}}{x_k^{c_i}}$ where $1 \leq i \leq r$, $\vec{v} \in \bz_{\geq 0}^{m + 1}$
with $\deg \vec v = c_i$. Furthermore, it has
relations generated by the following two forms:
\begin{enumerate}
	\item For each $(i, j)$ with $j \geq i + 2$ and each $\vec{v} \in \mss_i,
\vec{w} \in \mss_j$, there is a relation of the form: either
\begin{flalign*}
	&G_{i, j}^{\vec{v}, \vec{w}} = F_i^{\vec{v}} F_j^{\vec{w}}
	- \prod_{\vec{y} \in \mss_{h_{i, j}}} (F_{h_{i, j}}^{\vec{y}})
	^{g_{\vec{y}}} &(i < h_{i, j} < j) \\
	\text{or} \\
	&G_{i, j}^{\vec{v}, \vec{w}} = F_i^{\vec{v}} F_j^{\vec{w}}
	- \prod_{\vec{y} \in \mss_{h_{i, j}}} (F_{h_{i, j}}^{\vec{y}})^{g_{\vec{y}}}
	\cdot \prod_{ \vec{z} \in
	\mss_{h_{i, j} + 1}} (F_{k_{i, j} + 1}^{\vec{z}})^{g'_{\vec{z}}}
	&(i < h_{i, j} < h_{i, j} + 1 < j).
\end{flalign*}
\todo{Peter: Can you clarify when these two cases occur: i.e. does the first 
only occur if $j = h_{i,j}+1$; otherwise, in cases when $i< h_{i,j} < h_{i,j} + 1< 
j$, do we get relations of both forms? \\
Robin: It purely depends on whether the product $F_i^{\vec{v}} F_j^{\vec{w}} 
$ happens to lie on the ray given by $\frac{c_{h_{i, j}}}{d_{h_{i, j}}}$ which is 
not clearly determined by $i$ or $j$ from what I can see... Unless you see 
something else. Only one relation exists for each $(i, j)$ and it is exactly one 
of the two forms. I figured that the ``either'' made that clear, no?}
	\item For each $(i, j)$ with
$j = i$ or $j = i + 1$ and each $\vec{v} \in \mss_i, \vec{w} \in
\mss_j$ with $\vec{v} \not\prec \vec{w}$ (see Definition
~\ref{defn:vec-order}) there is a relation of the form
\begin{align*}
	&L_{i, j}^{\vec{v}, \vec{w}} = F_i^{\vec{v}} F_j^{\vec{w}}
	- F_i^{\vec{y}} F_j^{\vec{z}} \\
\end{align*}

@@

Next, we show that the relations given in the statement of the theorem generate all relations. 
In particular, if $j \geq i + 2$ then $F_i^{\vec{v}} F_j^{\vec{w}}$
has a decomposition of the form ~\eqref{eqn:one-point-canonical-form} of 
products of generators in adjacent degrees where $h$
depends on $i$ and $ j$, so we notate $h_{i, j} := h \in \{1, \ldots, r\}$. We 
also have that $i \leq h_{i, j} < j$ since $(d_i + d_j, c_i + c_j)$ is in the
$\bz_{\geq 0}$-span of $(d_{h_{i, j}}, c_{h_{i, j}})$ and
$(d_{h_{i, j} + 1}, c_{h_{i, j} + 1})$. Furthermore,
$h_{i, j} \neq i$ and $h_{i, j} \neq j - 1$ by the proof of the analogous 
statement in $bp^1$ given by O'Dorney
\cite[Theorem 6]{dorney:canonical}. This gives the relations
$G_{i, j}^{\vec{v}, \vec{w}}$.
